### 2025-10-12 thesis/design/functions.tex
Finish the design section for now

@@

The function analyzes the selected examples in $c$ to identify which schema elements
(tables, columns, foreign key relationships) are commonly referenced in similar queries.
This analysis enables the system to steer the model's attention on the most relevant
portions of potentially large and complex database schemas, thereby improving both accuracy
and computational efficiency and prevents exceeding limited context windows.

@@

@@

@@

The voting function $\nu$ implements a majority voting algorithm similar to that
described by OmniSQL \citep{OmniSQL}, where the result that appears most frequently
across multiple generation attempts is deemed to be the most likely correct answer.
The function applies frequency-based selection among the valid candidates. In cases where
no clear majority exists, the function may apply additional heuristics such as query
complexity or execution performance to determine the final proposed query candidate $w_{consensus}$.

>>>>

\begin{algorithm}
\caption{$\sigma$ - Example Selection}\label{algorithms:sigma}
\begin{algorithmic}[1]
\Require $q \in \mathcal{Q}$, $z \in \Omega_{\mathcal{S}}$, $s \in \mathcal{S}$, $v \in \mathcal{V}$, $d \in \mathcal{D}$
\Require $k \in \mathbb{N}$, $k \geq 1$                            \Comment{Number of examples to select}
\State $q' \gets mask(q)$                                          \Comment{Mask input question}
\State $z' \gets mask(z)$                                          \Comment{Mask zero-shot result}
\State $candidates  \gets \emptyset$                               \Comment{Initialize candidate set}
\For{$(q_i, s_i, \omega_i) \in v$}                            \Comment{For each sample}
    \State $sim_q \gets cosine(\iota(q'), \iota(mask(q_i)))$       \Comment{Calculate question similarity}
    \State $sim_s \gets cosine(\iota(z'), \iota(mask(\omega_i)))$  \Comment{Calculate SQL similarity}
    \State $sim_d \gets d(s, s_i)$                                 \Comment{Calculate schema distance}
    \State $score \gets w_q \cdot sim_q + w_s \cdot sim_s + w_d \cdot sim_d$
    \State $candidates \gets candidates \cup \{(q_i, s_i, \omega_i, score)\}$
\EndFor
\State \Return $top_{k}(candidates)$                             \Comment{Return k highest scoring examples}
\end{algorithmic}
\end{algorithm}

@@

The function analyzes the selected examples in $c$ to identify which schema elements
are commonly referenced in similar queries. This analysis enables the system to
steer the model's attention on the most relevant portions of potentially large
and complex database schemas, thereby improving both accuracy and computational
efficiency and prevents exceeding limited context windows.

\begin{algorithm}
\caption{$\phi$ - Schema Subsetting}\label{algorithms:phi}
\begin{algorithmic}[1]
\Require $c \in \mathcal{C}'$, $s \in \mathcal{S}$
\State $s' \gets \emptyset$                          \Comment{Empty schema subset}
\For{$table \in s$}
    \If{$\exists~r \in references(c, table)$}        \Comment{Any candidates reference the table}
        \State $s' \gets s' \cup \{table\}$          \Comment{Extend subsetted schema}
    \EndIf
\EndFor
\State \Return $s'$                                  \Comment{Return subsetted schema}
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
\caption{$\pi$ - Query Projection}\label{algorithms:pi}
\begin{algorithmic}[1]
\Require $q \in \mathcal{Q}$, $s \in \mathcal{S}$, $c \in \mathcal{C}'$
\Require $\tau \in \mathbb{R}$, $\tau > 0$                     \Comment{Relevance threshold}
\State $c_{rel} \gets \{~(q_i, s_i, \omega_i, d_i) \in c : d_i < \tau~\}$  \Comment{Filter by relevance}
\State $ex \gets \{~fmt(q_i, \omega_i, d_i)~|~(q_i, s_i, \omega_i, d_i) \in c_{rel}~\}$  \Comment{Format examples}
\State $prompt \gets prompt(q, s, ex)$                         \Comment{Construct prompt}
\State $out \gets model(prompt)$                               \Comment{Model inference}
\State $\Omega_{cand} \gets extract_{sql}(out)$                \Comment{Extract SQL candidates}
\For{$\omega_{raw} \in \Omega_{cand}$}                         \Comment{Validate candidates}
    \If{$\omega_{raw} \in \Omega_{\mathcal{S}}$}               \Comment{Check syntactic validity}
        \State \Return $\omega_{raw}$                          \Comment{Return first valid query}
    \EndIf
\EndFor
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
\caption{$\rho$ - Self Refinement}\label{algorithms:rho}
\begin{algorithmic}[1]
\Require $q \in \mathcal{Q}$, $s \in \mathcal{S}$, $e \in \mathcal{E}$, $\omega_{raw} \in \Omega_{\mathcal{S}}$
\State $exec \gets e(s, \omega_{raw})$                          \Comment{Verify execution output}
\State $prompt \gets prompt_{refine}(q, s, \omega_{raw}, exec)$ \Comment{Construct refinement prompt}
\State $out \gets model(prompt)$                                \Comment{Generate refinement output}
\State $\Omega_{ref} \gets extract_{sql}(out)$                  \Comment{Extract refined candidates}
\For{$\omega_{ref} \in \Omega_{ref}$}                           \Comment{Validate refined queries}
    \If{$\omega_{ref} \in \Omega_{\mathcal{S}}$}                \Comment{Check syntactic validity}
        \State \Return $\omega_{ref}$                           \Comment{Return first valid refinement}
    \EndIf
\EndFor
\end{algorithmic}
\end{algorithm}

@@

The voting function $\nu$ implements a majority voting algorithm similar to that
described by OmniSQL \citep{OmniSQL}, where the result that appears most frequently
across multiple generation attempts is deemed to be the most likely correct answer.
The function applies frequency-based selection among the valid candidates. In cases where
no clear majority exists, the function may apply additional heuristics such as query
complexity or execution performance to determine the final proposed query candidate $\omega_{consensus}$.

\begin{algorithm}
\caption{$\nu$ - Consensus Voting}\label{algorithms:nu}
\begin{algorithmic}[1]
\Require $C \in \mathcal{C}$, $e \in \mathcal{E}$
\State $results \gets \{\}$                               \Comment{Map from result sets to candidate lists}
\For{$\omega \in C$}                                      \Comment{Execute each candidate}
    \State $result \gets e(\omega)$                       \Comment{Execute query}
    \State $results[result] \gets results[result] \cup \{\omega\}$
\EndFor
\State $dist \gets \{results[r] : r \in results\}$        \Comment{Get all result groups}
\State $dist \gets sort(dist, |group| \mapsto |group|)$   \Comment{Sort by group size (largest first)}
\State \Return $top_1(top_1(dist))$                       \Comment{Return largest group}
\end{algorithmic}
\end{algorithm}
